---
output: github_document
always_allow_html: true

---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```
The bangladesh package provides ready-to-use shapefiles for different administrative 
regions of Bangladesh (e.g., Division, District, Upazila, and Union). 
Usually, it is difficult to plot choropleth maps for Bangladesh in R. 
This package will help users to draw thematic maps of administrative regions 
of Bangladesh easily as it comes with the sf objects for the boundaries and 
regions’ names in English. It also provides functions allowing users to efficiently
get specific area maps and center coordinates for regions. Users can also search for
a specific area and calculate the centroids of those areas.


## Getting Started

This packages comes with sf objects for administrative levels 0-4 (Country, 
Division, District, Upazila, Union). The easiest way to get the shapefile for
a level is to is to use `get_map()` function.

```{r setup}
# remotes::install_github("ovirahman/bangladesh")
library(bangladesh)

country <- get_map("country")
division <- get_map("division")
district <- get_map("district")
upazila <- get_map("upazila")
union <- get_map("union")
```

## Plotting Map

To start with we can check the sample function `bd_plot()` to draw the map of different 
administrative levels of Bangladesh, which uses [`tmap`](https://cran.r-project.org/package=tmap), a very flexible and cool package to visualize thematic maps.

```{r plot, warning=FALSE, message=FALSE,fig.width = 5, fig.height = 5, fig.align = "center"}
bd_plot("country")
bd_plot("division")
bd_plot("district")
```

We can also plot beautiful interactive maps with this.

```{r interactive, include = FALSE, warning=FALSE, message=FALSE,fig.width = 5, fig.height = 5, fig.align = "center"}
bd_plot(level = "district", type = "interactive")
```


## Choropleth map with data 

Using the `tmap` package (my favorite for creating thematic maps), we can 
make cool choropleths, both static and interactive. When plotting mode is chosen as static (plot) it 
returns a `ggplot` object, when interactive (view) it returns a [`leaflet`](https://cran.r-project.org/package=leaflet) object. 

```{r choropleth_static, warning=FALSE, message=FALSE,fig.width = 5, fig.height = 5, fig.align = "center"}
library(tmap)
population <- bangladesh::pop_district_2011[, c("district", "population")]
district <- get_map("district")

map_data <- dplyr::left_join(district, population, by = c("District" = "district"))

map <- tm_shape(map_data) + 
  tm_polygons("population",id = "District",palette = "Reds", title = "Population") +
  tm_style("cobalt")+
  tm_layout(
    "Bangladesh District Wise Population Map\nSource: BBS",
    title.position = c("left", "bottom"),
    legend.position = c("right", "top")
    )

tmap::tmap_mode("plot")
map
```
To make it interactive we can do the following:

```{r choropleth_interactive, include = FALSE, warning=FALSE, message=FALSE,fig.width = 5, fig.height = 5, fig.align = "center"}
tmap::tmap_mode("view")
map
```


## Using ggplot2 and leaflet

We can also use `ggplot2` and `leaflet` to draw customized choropleths with the sf objects provided in
`bangladesh` package.


```{r ggplot, warning=FALSE, message=FALSE,fig.width = 5, fig.height = 5, fig.align = "center"}
library(ggplot2)
ggplot(data = map_data) +
  geom_sf(aes(fill = population))+
  theme_void()+
  viridis::scale_fill_viridis(trans = "log", name="Population", labels = scales::unit_format(unit = "M", scale = 1e-6)) +
  labs(
    title = "Bangladesh Population Map",
    subtitle = "Population & Housing Census 2011",
    caption = "Data Source: BBS"
  )
  
```

## Other useful functions

It is also possible to get the approximate center points (centroids) of administrative regions easily 

by using `get_coordinates()` function in `bangladesh` package.


```{r centroids, warning=FALSE, message=FALSE,fig.width = 5, fig.height = 5, fig.align = "center"}
division_map <- get_map("division")
division_centroids <- bangladesh::get_coordinates(level = "division")
knitr::kable(division_centroids, format = "html")
ggplot(data = division_map) +
  geom_sf() +
  geom_sf_label(aes(label = Division)) +
  geom_point(data = division_centroids, x = division_centroids$lon, y = division_centroids$lat, col = "red", size = 3) +
  xlab("")+ ylab("")+
  theme_minimal()


```


Suppose someone needs to plot partially  a single or selected number of divisions 
instead of whole country map, in that case the function `get_divisions()` might be beneficial. 

```{r partial_map, warning=FALSE, message=FALSE,fig.width = 5, fig.height = 5, fig.align = "center"}

sylhet <- get_divisions(divisions = "Sylhet",level =  "upazila")
# single division
ggplot(data = sylhet) +
  geom_sf() +
  xlab("")+ ylab("")+
  theme_minimal()

#multiple division
sylhet_chittagong_dhaka <- get_divisions(divisions = c("Sylhet", "Chittagong", "Dhaka"),level =  "upazila")
ggplot(data = sylhet_chittagong_dhaka) +
  geom_sf() +
  xlab("")+ ylab("")+
  theme_minimal()


```

To search for an area within the provided names for administrative regions we can
apply the `bd_search()` function. The result can also include centroids for those areas.


```{r search, warning=FALSE, message=FALSE,fig.width = 5, fig.height = 5, fig.align = "center"}
amtali <- bd_search("amtali", level = "union", as.is = T, coordinates = T)
knitr::kable(amtali, format = "html")
ggplot(bangladesh::map_union) +
  geom_sf() +
  geom_point(data = amtali, x = amtali$lon, y = amtali$lat, col = "red", size = 3) 


```
